---
title: "Cancer Rates Across the U.S. by aem303"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    includes: 
        after_body: "busy.html"
runtime: shiny
---

<style type="text/css">

    .storyboard-nav .sbframelist {
        margin: 0 auto;
        width: 94%;
        height: 70px;
        overflow: hidden;
        text-shadow: none;
        margin-bottom: 8px;
    }

    .storyboard-nav .sbnext, .storyboard-nav .sbprev {
        float: left;
        width: 2%;
        height: 50px;
        font-size: 50px;
    }



</style>

```{r setup, include = FALSE}
library(leaflet)
library(shiny)
library(shinyjs)
library(flexdashboard)
library(plotly)
library(shinycssloaders)


map_data <- readRDS('map_data.rds')
map_data_df <- as.data.frame(map_data)
map_data_df <- map_data_df %>% mutate(loc = paste0(county,", ",state))
popup_dat <- readRDS('popup_cancer.rds')
bins_sites <- c(0, 1, 2, 3, 5, 8, 10, 15, 20, Inf)
pal_sites <- colorBin("YlOrRd", domain = map_data$num_sites, bins = bins_sites)
bins_pp <- c(0, 1,2,3, 5, 12, 25, 36, 51, Inf)
pal_pp <- colorBin("YlOrRd", domain = map_data$num_pp, bins = bins_pp)
#pal_pp <- colorQuantile("YlOrRd", NULL, n = 4)

pal_cancer <- colorQuantile("YlOrRd", NULL, n = 9)

react_list <- reactiveValues()
react_list$num_traces_plotly_cancer = 0
react_list$num = 25
k <- 3

```


### Distribution of Cancer Rates

```{r}
plotlyOutput("plotly_cancer_dist")
output$plotly_cancer_dist <- renderPlotly({
  map_data_df_dist <- map_data_df %>% filter(!state %in% input$exclude_state_dist_us)
  plot_ly(
          x = map_data_df_dist[[input$dist_level_us]], 
          y = map_data_df_dist[['cancer_per100k']], 
          color = map_data_df_dist[[input$dist_level_us]], 
          type = "box", 
          boxpoints = "all", 
          jitter = 0.3,
        pointpos = -1.0,
        hoverinfo = "text",
                      hovertext = paste0("<b>Location: </b>", 
                          map_data_df_dist$county,", ",map_data_df_dist$state, 
                          "<br><b>Cancer/100k: </b>", 
                          map_data_df_dist$cancer_per100k))
        })

# data = map_data_df %>% mutate(cancer_pCI = cancer_95ciH - cancer_per100k, cancer_mCI = cancer_per100k - cancer_95ciL) %>% filter(!is.na(cancer_per100k)) %>% arrange(cancer_per100k)
# data <- data %>% mutate(order = 1:nrow(data))
# plot_ly(data = data,
#         x = ~order,
#         y = ~cancer_per100k,
#         error_y = list(array = ~cancer_pCI),
#         type = 'scatter',
#         mode = 'markers'
# )

```

***

The state by state distribution of counties shows clear differences in cancer rates.

```{r}

selectInput("dist_level_us","Select U.S. Breakdown Type",list("Census Region (4)" = "region",
                                                           "Census Division (9)" = "division",
                                                           "States" = "state"))

pickerInput("exclude_state_dist_us","Exclude U.S. States",choices = c('AL','AK','AZ','AR','CA','CO','CT','DC','FL','GA','HI','ID','IL','IN','IA','KY','LA','ME','MD','MA','MI','MS','MO','MT','NE','NJ','NM','NY','NC','ND','OH','OK','OR','PA','SC','SD','TN','TX','UT','VA','WA','WV','WI','WY','DE','NV','NH','VT','RI'), options = list(`actions-box` = TRUE),multiple = T)
```


### Intro: Heatmap of Cancer Rates by County {data-commentary-width=375}

```{r, eval = TRUE}




fillRow(leafletOutput("intro_map"))


output$intro_map <- renderLeaflet({
  leaflet(data = map_data) %>%
  # Base groups
  addTiles() %>%
  setView(lng = -105, lat = 40, zoom = 4) %>% 
  addPolygons(fillColor = ~pal_cancer(cancer_per100k), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1,
              layerId = ~GEOID,
              group="Cancer Rate/100,000 by Counties")
})



# observeEvent(input$intro_map_click, {
# 
#     click <- input$intro_map_click
#     text<-paste("Latitude ", round(click$lat,2), "Longtitude ", round(click$lng,2))
# 
#     proxy <- leafletProxy("intro_map")
# 
#     ## This displays the pin drop circle
#     proxy %>% 
#         clearGroup("new_point") %>%
#         #clearMarkers(layerId=input$mymap_click$id) %>%
#         #addPopups(click$lng, click$lat) %>%
#         addCircles(click$lng, click$lat, radius=30000, color="blue", group = "new_point")
# 
# })

observeEvent(input$go_worst10, {
    
    worst_5 <- map_data_df %>% dplyr::arrange(desc(cancer_per100k)) %>% head(react_list$num)
    worst_5['order'] <- 1:nrow(worst_5)
    #text<-paste("Latitude ", round(worst_5$lat,2), "Longtitude ", round(worst_5$lng,2))

    proxy <- leafletProxy("intro_map")

    ## This displays the pin drop circle
    proxy %>% 
        clearGroup("worst_points") %>%
        addCircles(worst_5$longitude, worst_5$latitude, radius=50000, color="blue", group = "worst_points", 
                   popup = paste0("<strong>County: </strong>", 
                                                    worst_5$county, 
                                                    "<br><strong>Cancer/100k: </strong>", 
                                                    worst_5$cancer_per100k,
                                                    "<br><strong>Rank: </strong> ",worst_5$order))

})

observeEvent(input$go_best10, {
    
    best5 <- map_data_df %>% dplyr::arrange(cancer_per100k) %>% head(react_list$num)
    best5['order'] <- 1:nrow(best5)
    #text<-paste("Latitude ", round(worst_5$lat,2), "Longtitude ", round(worst_5$lng,2))

    proxy <- leafletProxy("intro_map")

    ## This displays the pin drop circle
    proxy %>% 
        clearGroup("best_points") %>%
        addCircles(best5$longitude, best5$latitude, radius=50000, color="green", group = "best_points", 
                   popup = paste0("<strong>County: </strong>", 
                                                    best5$county, 
                                                    "<br><strong>Cancer/100k: </strong>", 
                                                    best5$cancer_per100k,
                                                    "<br><strong>Rank: </strong> ",best5$order))

})
radius = 3
# When circle is clicked on...show a popup
observeEvent(input$intro_map_shape_click$id, {
    pointId <- input$intro_map_shape_click$id
    shape_i = map_data_df %>% filter(GEOID == pointId)
    offset = isolate((input$intro_map_bounds$north - input$intro_map_bounds$south) / (23 + radius + (18 - input$intro_map_zoom)^2 ))

    leafletProxy("intro_map") %>% addPopups(lat = shape_i$latitude + offset, lng = shape_i$longitude, 
                                                    paste0("<strong>County: </strong>", 
                                                    shape_i$county, 
                                                    "<br><strong>Cancer/100k: </strong>", 
                                                    shape_i$cancer_per100k))
  })


observeEvent(input$best_popup, {
    leafletProxy("intro_map") %>%
    clearPopups()
    best <- map_data_df %>% dplyr::arrange(desc(cancer_per100k)) %>% head(react_list$num)
    best['order'] <- 1:nrow(best)
    offset = isolate((input$intro_map_bounds$north - input$intro_map_bounds$south) / (23 + radius*1.8 + (18 - input$intro_map_zoom)^2 ))

    #shape_i <- best[react_list$num,]
    leafletProxy("intro_map") %>% addPopups(lat = best$latitude + offset, lng = best$longitude, 
                                                  paste0("<strong>County: </strong>", 
                                                  best$county, 
                                                  "<br><strong>Cancer/100k: </strong>", 
                                                  best$cancer_per100k,
                                                    "<br><strong>Rank: </strong> ",best$order))

    #react_list$num <- react_list$num + 1
  })

observeEvent(input$worst_popup, {
    leafletProxy("intro_map") %>%
    clearPopups()
    worst <- map_data_df %>% dplyr::arrange(cancer_per100k) %>% head(react_list$num)
    worst['order'] <- 1:nrow(worst)
    offset = isolate((input$intro_map_bounds$north - input$intro_map_bounds$south) / (23 + radius*1.8 + (18 - input$intro_map_zoom)^2 ))

    #shape_i <- best[react_list$num,]
    leafletProxy("intro_map") %>% addPopups(lat = worst$latitude + offset, lng = worst$longitude, 
                                                  paste0("<strong>County: </strong>", 
                                                  worst$county, 
                                                  "<br><strong>Cancer/100k: </strong>", 
                                                  worst$cancer_per100k,
                                                    "<br><strong>Rank: </strong> ",worst$order))

    #react_list$num <- react_list$num + 1
  })
```


```{r}
# TRIED TO GET POPUPS TO COME UP ONE AT A TIME
# observe({
#     req(!is.null(react_list$num))
#     req(react_list$num <= 10 & react_list$num >= 2)
#     
#     best <- map_data_df %>% dplyr::arrange(desc(cancer_per100k)) %>% head(num)
#       offset = isolate((input$intro_map_bounds$north - input$intro_map_bounds$south) / (23 + radius*1.5 + (18 - input$intro_map_zoom)^2 ))
#   
#       shape_i <- best[react_list$num,]
#       leafletProxy("intro_map") %>% addPopups(lat = shape_i$latitude + offset, lng = shape_i$longitude, 
#                                                     paste0(shape_i$county, 
#                                                     ": ", 
#                                                     shape_i$cancer_per100k))
#       
#       react_list$num <- react_list$num + 1
#       
# })

observeEvent(input$clear_intro_map,{
  leafletProxy("intro_map") %>%
    clearPopups()
  leafletProxy("intro_map")  %>% 
        clearGroup("best_points")
  leafletProxy("intro_map")  %>% 
        clearGroup("worst_points")
  react_list$num <- input$num_intro

})
```

***

This project investigates county-level cancer rates around the United States. There are many factors that contribute to cancer, including hereditary and environmental causes. This story will focus on how environmental contamination from toxic waste sites and powerplants have an effect on cancer rates. 

Start off by ***clicking*** on the heatmap showing cancer rates in a few counties to get yourself familiar with the NIH data for age-adjusted cancer rates. ***Clicking*** a county will show its name and its age-adjusted cancer rate per 100 thousand people. In subsequent frames more variables will be introduced.

For now, let’s examine the distribution of cancer rates around the country. Click the buttons below to see the location of the 25 highest and 25 lowest cancer rate counties are located on the map. Click on the info buttons to show info for the highest or lowest counties in the US.

```{r}
fluidRow(actionButton("go_worst10", "Highest rate counties"),
actionButton("go_best10", "Lowest rate counties"),align="center")
```

```{r}
fluidRow(actionButton("best_popup", "Highest counties info"),
         actionButton("worst_popup", "Lowest counties info"),align="center")
```

Adjust the # below to show greater or fewer counties in the rankings. You must reset the map after changing the value.

```{r}
fluidRow(numericInput("num_intro", NULL, value = 25, min = 1, max = 100, step = 5, width = '25%'),align="center")
```

You must reset the map to make changes to the number of items to show in the ranking.

```{r}
fluidRow(actionButton("clear_intro_map", "Reset Map"),align="center")
```

The West and Southwest have many of the counties with the lowest cancer rates. Many of the higher cancer rate counties are in the South, particularly in Kentucky. The most interesting border between the best and worst counties appears near the tri-cities region in southwest Virginia. Kentucky counties are experiencing the highest cancer rates, while their Virginian neighbors have the lowest rates in the country.

### Heatmap of Cancer Rates by County <br> East Coast Problems {data-commentary-width=600}

```{r, eval = TRUE}
leafletOutput("cancer_map")


output$cancer_map <- renderLeaflet({
  leaflet(data = map_data) %>%
  # Base groups
  addTiles() %>%
  setView(lng = -105, lat = 40, zoom = 4) %>% 
  addPolygons(fillColor = ~pal_cancer(cancer_per100k), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1,
              popup = popup_dat,
              layerId = ~GEOID,
              group="Cancer Rate/100,000 by Counties")
})

```

*** 

```{r}
# A reactive expression that returns the set of counties that are
# in bounds right now
dataInBounds <- reactive({
  if (is.null(input$cancer_map_bounds))
    return(map_data[FALSE,])
  bounds <- input$cancer_map_bounds
  latRng <- range(bounds$north, bounds$south)
  lngRng <- range(bounds$east, bounds$west)

  subset(map_data,
    latitude >= latRng[1] & latitude <= latRng[2] &
      longitude >= lngRng[1] & longitude <= lngRng[2])
})

observeEvent(input$cancer_map_shape_click, { 
 
    if (react_list$num_traces_plotly_cancer == 1){
      plotlyProxy("cancer_plotly", session) %>%
        plotlyProxyInvoke("deleteTraces", 1)
    }
    react_list$num_traces_plotly_cancer <- 1

  
  print(input$cancer_map_shape_click$id)
  df_cancer_sub <- react_list$df_cancer %>% filter(GEOID %in% c(input$cancer_map_shape_click$id))
  
  print(df_cancer_sub$trend5yr)
  print(df_cancer_sub$cancer_per100k)
  print(react_list$num_traces_plotly_cancer)
  
  plotlyProxy("cancer_plotly", session) %>%
  plotlyProxyInvoke("addTraces", list( 
                                    x = list(df_cancer_sub$trend5yr,NA),
                                    y = list(df_cancer_sub$cancer_per100k,NA),
                                    type = 'scatter',
                                    mode = 'markers',
                                    name = paste0(df_cancer_sub$county,', ',df_cancer_sub$state)[1],
                                    marker = list(size = 12.5,
                                                  color = 'rgba(255,51,51,0.8)',
                                                  line = list(color = 'rgba(255, 0, 0, .7)',
                                                            width = 2))
  )
  )                     
})


observeEvent({event_data("plotly_click", source="cancer_plotly")}, {
    # get clicked point
    click_data <- event_data("plotly_click", source="cancer_plotly")
    test_df <<- react_list$df_cancer
    shape_i = react_list$df_cancer[click_data[['pointNumber']] + 1,]
    offset = isolate((input$cancer_map_bounds$north - input$cancer_map_bounds$south) / (23 + 7 + (18 - input$cancer_map_zoom)^2 ))
    
    message("Plotly point clicked:")
    message(shape_i$county)
    message(click_data)
    
    leafletProxy("cancer_map") %>% addPopups(lat = shape_i$latitude + offset, lng = shape_i$longitude, 
                                             paste0("<strong>County: </strong>", 
                                                    shape_i$county, 
                                                    "<br><strong>Cancer/100k: </strong>", 
                                                    shape_i$cancer_per100k))
  })


# # SEE PLOTLY CLICK DATA
# renderPrint({
#     message('clicked')
#     d <- event_data("plotly_click", source="cancer_plotly")
#     print(d)
#     if (!is.null(d)) d
#   })
#   
# renderPrint({
#   d <- event_data("plotly_click", source="cancer_plotly")
#   if (!is.null(d)){
#     shape_i = react_list$df_cancer[d[['pointNumber']] + 1,]
#     shape_i$county
#     message(shape_i$county)
#     message(d)
#   }
# })


observeEvent(input$clear_cancer_popup_button, {
    leafletProxy("cancer_map") %>% clearPopups()
})


plotlyOutput("cancer_plotly")
output$cancer_plotly <- renderPlotly({
  react_list$df_cancer_map <- dataInBounds()
  react_list$df_cancer <- as.data.frame(react_list$df_cancer_map) %>% arrange(GEOID) %>% filter(!is.na(cancer_per100k)) %>% filter(!is.na(trend5yr))
  plot_ly(data = react_list$df_cancer,
                      x = ~trend5yr, 
                      y = ~cancer_per100k,
                      type = "scatter",
                      mode = "markers",
                      source="cancer_plotly",
                      hoverinfo = "text",
                      hovertext = paste0("<b>Location: </b>", 
                          react_list$df_cancer$county,", ",react_list$df_cancer$state, 
                          "<br><b>Cancer/100k: </b>", 
                          react_list$df_cancer$cancer_per100k)
                      )
  })

# popup_SF <- paste0("<strong>Superfund Name: </strong>", 
#                    df_fips$park, 
#                    "<br><strong>Type: </strong>", 
#                    natparks.points$type)
actionButton("clear_cancer_popup_button", "Clear popups on map")
```

Some commentary about Frame 1.

### Frame 2 {data-commentary-width=400}

***

#### Plot1: wt, hp

```{r}
plotDf <- mtcars[,c("wt","hp")]
plot(plotDf)
```


```{r}
plotDf <- mtcars[,c("mpg","disp")]
plot(plotDf)
```

https://shiny.rstudio.com/gallery/superzip-example.html 

Some commentary about Frame 2.

### Frame 3

```{r echo=FALSE}

library(DT)

DT::dataTableOutput('irisTable')
output$irisTable = DT::renderDataTable(iris, selection = 'multiple')

```

***
```{r}


p("By default DT allows multiple row selection. Selected rows are...")
renderPrint(input$irisTable_rows_selected)
```

### Frame 4

More specific commentary 

use leaflet proxy to select specific points on the map that are the worst for cancer

https://stackoverflow.com/questions/48781380/shiny-how-to-highlight-an-object-on-a-leaflet-map-when-selecting-a-record-in-a

### Frame 5

Run models to see the most important sites. have to have results already made.

### Frame 6
Regression results showing prediction and actual relationship

### Frame 7 conclusion, references, and github.
