---
title: "Storyboard Commentary"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
runtime: shiny
---

### Frame 0

INTRODUCE TOPIC, SHOW SOME STATS

### Frame 1

SHOW MAP OF CANCER USING LEAFLET

```{r}
renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser Eruption Duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

*** 
```{r}
selectInput("n_breaks", label = "Number of bins:",
            choices = c(10, 20, 35, 50), selected = 20)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
            min = 0.2, max = 2, value = 1, step = 0.2)
```

Some commentary about Frame 1.

### Frame 2 {data-commentary-width=400}

SHOW DOTS AND COLORS BASED ON NUMBER OF SUPERFUND SITES

POINT OUT SPECIFIC SITES THAT ARE BAD

#### Plot1: wt, hp

```{r}
plotDf <- mtcars[,c("wt","hp")]
plot(plotDf)
```


```{r}
plotDf <- mtcars[,c("mpg","disp")]
plot(plotDf)
```
***

https://shiny.rstudio.com/gallery/superzip-example.html 

Some commentary about Frame 2.

### Frame 3

MAP WITH POWERPLANTS

### Frame 4

More specific commentary 

use leaflet proxy to select specific points on the map that are the worst for cancer

https://stackoverflow.com/questions/48781380/shiny-how-to-highlight-an-object-on-a-leaflet-map-when-selecting-a-record-in-a

### Frame 5

Run models to see the most important sites. have to have results already made.

### Frame 6
Regression results showing prediction and actual relationship

### Frame 7 conclusion, references, and github.